Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-10-05 paper/article_jos.qmd
typos

The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$, can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend (sustained and systematic variations over a long period of time) and cycle (cyclical movements superimposed upon the long-term trend) components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (@sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], a usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\widehat{TC}_t$ and it can be shown that this equivalent to apply of moving average (@sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

For symmetric filters, the phase shift function is equal to zero (modulo $\pi$).
So there is no delay in detecting turning points.
However, they cannot be used in the beginning and in the end of the time series because no past or future value can be used. 
Thus, for real-time estimation, it is needed to build asymmetric moving averages that approximate the symmetric moving average.

@@

Since $d+1$ parameters are estimated with $\boldsymbol \beta$, we need at least $2h+1$ observations (thus $2h\geq d$) and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
$$
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
$$
where $\kappa_j$ is a set of weights called a *kernel*.
We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\transp{\boldsymbol X}\boldsymbol K\boldsymbol y.$
With $\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}$, the estimate of the trend is:
$$
\hat{m}_{t}=\widehat{TC}_t=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
$$ {#eq-mmsym}
To conclude, the estimate of the trend $\widehat{TC}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
Moreover, since $\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}$ we have:
$$
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\{1,2,\dots,d\right\}:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
$$
Hence, the filter $\boldsymbol w$ preserves deterministic polynomials of order $d$.

@@

```{r, echo=FALSE, out.width="90%"}
#| label: fig-mmpenteconcac-ex
#| fig-cap: "Comparison of the estimates of $|\\delta_1/\\sigma|$ with a local estimator and global estimator (with IC ratio) on a simulated series, the vertical lines being the simulated turning points."
#| fig-subcap: 
#|   - "Symmetric moving average used for $\\hat \\delta_1$."
#|   - "Asymmetric moving average using two future points ($q=2$) used for $\\hat \\delta_1$."
#| layout-ncol: 1
knitr::include_graphics(c("img/filters_used/mm_penteconcavite_ex.pdf",
                          "img/filters_used/mm_penteconcavite_ex2.pdf"))
```

>>>>

The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$, can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend (sustained and systematic variations over a long period of time) and cycle (cyclical movements superimposed upon the long-term trend) components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (@sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], a usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_d)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\widehat{TC}_t$ and it can be shown that this equivalent to apply of moving average (@sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

For symmetric filters, the phase shift function is equal to 0 (modulo $\pi$).
So there is no delay in detecting turning points.
However, they cannot be used in the beginning and in the end of the time series because no past or future value can be used. 
Thus, for real-time estimation, it is needed to build asymmetric moving averages that approximate the symmetric moving average.

@@

Since $d+1$ parameters are estimated with $\boldsymbol \beta$, we need at least $2h+1$ observations (thus $2h\geq d$) and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
$$
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
$$
where $\kappa_j$ is a set of weights called a *kernel*.
We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\transp{\boldsymbol X}\boldsymbol K\boldsymbol y.$
With $\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}$, the estimate of the trend is:
$$
\hat{m}_{t}=\widehat{TC}_t=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
$$ {#eq-mmsym}
To conclude, the estimate of the trend $\widehat{TC}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
Moreover, since $\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}$ we have:
$$
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\{1,2,\dots,d\right\}:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
$$
Hence, the filter $\boldsymbol \theta$ preserves deterministic polynomials of order $d$.

@@

```{r, echo=FALSE, out.width="90%"}
#| label: fig-mmpenteconcac-ex
#| fig-cap: "Comparison of the estimates of $|\\delta_1/\\sigma|$ with a local estimator and global estimator (with IC ratio) on a simulated series, the vertical lines being the simulated turning points."
#| fig-subcap: 
#|   - "Symmetric moving average used for $\\hat \\delta_1$."
#|   - "Asymmetric moving average using two future points ($q=2$) for $\\hat \\delta_1$."
#| layout-ncol: 1
knitr::include_graphics(c("img/filters_used/mm_penteconcavite_ex.pdf",
                          "img/filters_used/mm_penteconcavite_ex2.pdf"))
```
